### 2018-04-02 0_intro.tex
Add some words on control-flow encoding.

Traditionally, studies related to concurrent programming concern on more fundamental theoretical questions of designing race-free and lock-free parallel algorithms, asynchronous data structures and synchronisation primitives of a programming language. Unfortunately,
when it comes to 
%in cases of 
the real-world concurrent programs, the algorithmic level of abstraction is not enough for guaranteeing their properties of correctness and reliability. The reasons of this fact lie in the code optimisations that both compiler and hardware perform in order to increase performance as much as possible. For instance, Figure~\ref{simple_wmm_x86} provides simple example of
unexpected state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' reachable in x86 machines (such little examples that illustrate specific behaviours of a WMM are called \textit{litmus tests}).
%reordering of memory access instructions within single process allowed by the x86-TSO weak memory model, which potentially breaks the program logic. 
This behaviour is caused by write buffers used by all processors in x86 architecture. These buffers cache writes to shared variables, so that the writes to shared memory does not become visible by other processes immediately. In the example, the write `\texttt{MOV~[x],1}' performed in the process \texttt{P0} stores value~\texttt{1} into the shared variable~\texttt{[x]} in the write buffer of process~\texttt{P0}. Meanwhile, the write cache of the process \texttt{P1} may not have updated version of the variable~\texttt{[x]}, neither may have the main memory, so that the read `\texttt{MOV~EBX,[x]}' performed in the process~\texttt{P1} may read the initial value~\texttt{0}, even if this variable has been already updated in another thread. More presice description of x86-TSO memory model is given in Chapter~\ref{ch:wmm:x86}.

@@

Although weak memory studies is rather young research area, there exist frameworks and tools for exploring WMMs and examining simple programs with respect to the them. The state-of-the-art tool is \texttt{diy} (for \textit{do it yourself}), developed by the researchers from INRIA institute, France, and University of Cambridge, UK and firstly released back in 2010. It is the sofware suite for designing and testing weak memory models, it consists of the litmus tests generators \texttt{diy7}, \texttt{diycross7} and \texttt{diyone7}, the litmus tests concrete executor \texttt{litmus7} that runs tests on a physical machine while collecting its behaviours, and the weak memory models simulator \texttt{herd7} that implements reachability analysis for exploring states reachable under specified~WMM.

@@

The first tool that performs the WMM-awared portability analysis is \texttt{PORTHOS} introduced in April~2017~\cite{Porthos17}. This tool reduces described problem to the bounded model checking (BMC) problem, which can be solved via SMT-solver. This approach allows to capture symbolically the semantics of analysing program and both weak memory models into single SMT-formula. As most modern SMT-solvers are efficient enough to be able to operate the state space of size .... (\textbf{TODO}), the used method can be applicable in solving the real-world problems.

@@

Current work aims to improve the proof-of-concept tool \texttt{PORTHOS} firstly introduced in April 2017 in the work~\cite{Porthos17} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, reliability and mantainability.

>>>>

Traditionally, studies related to concurrent programming concern on more fundamental theoretical questions of designing race-free and lock-free parallel algorithms, asynchronous data structures and synchronisation primitives of a programming language. Unfortunately,
when it comes to 
%in cases of 
the real-world concurrent programs, the algorithmic level of abstraction is not enough for guaranteeing their properties of correctness and reliability. The reasons of this fact lie in the code optimisations that both compiler and hardware perform in order to increase performance as much as possible. For instance, Figure~\ref{simple_wmm_x86} provides simple example of
unexpected state `\texttt{(0:EAX=0~/\textbackslash~1:EAX=0)}' reachable on x86 machines (such little examples that illustrate specific behaviour of a WMM are called \textit{litmus tests}).
%reordering of memory access instructions within single process allowed by the x86-TSO weak memory model, which potentially breaks the program logic. 
This state is allowed because in x86 architecture each processor may cache the write to shared memory variable into its local write buffer, so that they do not become visible by other processes immediately. In the example, the write `\texttt{MOV~[x],1}' performed by process \texttt{P0} stores value~\texttt{1} to the shared variable~\texttt{[x]} into the write buffer of process~\texttt{P0}. Meanwhile, the write cache of the process \texttt{P1} may not have updated version of the variable~\texttt{[x]}, neither may have the main memory, so that the read `\texttt{MOV~EBX,[x]}' performed in the process~\texttt{P1} may read the initial value~\texttt{0}, even if this variable has been already updated in another thread. More presice description of x86-TSO memory model is given in Chapter~\ref{ch:wmm:x86}.

@@

Although weak memory studies is rather young research area, there exist frameworks and tools for exploring WMMs and examining simple programs with respect to the them. The state-of-the-art tool is \tool{diy} (for \textit{do it yourself}), developed by the researchers from INRIA institute, France, and University of Cambridge, UK and firstly released back in 2010. It is the sofware suite for designing and testing weak memory models, it consists of the litmus tests generators \tool{diy7}, \tool{diycross7} and \tool{diyone7}, the litmus tests concrete executor \tool{litmus7} that runs tests on a physical machine while collecting its behaviours, and the weak memory models simulator \tool{herd7} that implements reachability analysis for exploring states reachable under specified~WMM.

@@

The first tool that performs the WMM-awared portability analysis is \tool{PORTHOS} introduced in April~2017~\cite{Porthos17}. This tool reduces described problem to the bounded model checking (BMC) problem, which can be solved via SMT-solver. This approach allows to capture symbolically the semantics of analysing program and both weak memory models into single SMT-formula. As most modern SMT-solvers are efficient enough to be able to operate the state space of size .... (\textbf{TODO}), the used method can be applicable in solving the real-world problems.

@@

Current work aims to rework the proof-of-concept tool \tool{PORTHOS} by extending the input language, which currently represents the minimum subset of C, and revising the general architecture of the tool in order to enhance performance, reliability and mantainability.
%interprocedural analysis
We called the new tool \textbf{the \tool{mousquetaires} framework} as now it not only performs the portability analysis, but can serve as the basis for SMT-solver driven weak memory model analysis.

### 2018-04-02 1_2_encod.tex
Cleanup wmm intro part. Set up named references. Fix too many commas error. Fix other citing errors.
Add some words on control-flow encoding.

\begin{definition}[Portability]
Let $\mathcal{M_S}$, $\mathcal{M_T}$ be two weak memory models. A program $P$ is portable from $\mathcal{M_S}$ to $\mathcal{M_T}$ if 
$cons_{\mathcal{M_T}}(P) \subseteq cons{\mathcal{M_S}}(P)$
\end{definition}

@@

It is possible to formulate this requirement as an SMT formula, so that the portability analysis problem becomes reduced to the BMC problem. The full SMT formula $\phi$ should contain encodings of control-flow ($\phi_{CF}$) and data-flow ($\phi_{DF}$) of the program, and assertions of both memory models: $\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}$. The control-flow and data-flow encodings are standard for BMC~\cite{collavizza2006exploration}, they are described below. However, encoding of memory models requires additional techniques due to recursive definitions of relations, that were proposed in~\cite{Porthos17}.
%$The \phi_{\mathcal{M_T}}$ is an encoding of the derived relations and all assertions of
%, and φ ¬M S

\subsection{Encoding of the control-flow constraints}
\label{ch:port:enc:cf}

\subsection{Encoding of the data-flow constraints}
\label{ch:port:enc:cf}

\subsection{Encoding of the memory model constraints}
\label{ch:port:enc:wmm}

>>>>

\begin{definition}[Portability]
Let $\mathcal{M_S}$, $\mathcal{M_T}$ be two weak memory models. A program $P$ is portable from $\mathcal{M_S}$ to $\mathcal{M_T}$ if 
$cons_{\mathcal{M_T}}(P) \subseteq cons_{\mathcal{M_S}}(P)$
\end{definition}

@@

It is possible to formulate this requirement as an SMT formula, so that the portability analysis problem becomes reduced to the BMC problem. The full SMT formula $\phi$ should contain encodings of control-flow ($\phi_{CF}$) and data-flow ($\phi_{DF}$) of the program, and assertions of both memory models: $\phi = \phi_{CF} \land \phi_{DF} \land \phi_{\mathcal{M_T}} \land \phi_{\lnot\mathcal{M_S}}$. If the formula is satisfiable, there exist a portability bug.
%The control-flow and data-flow encodings are standard for BMC~\cite{collavizza2006exploration}, they are described below. 
%However, encoding of memory models requires additional techniques due to recursive definitions of relations, that were proposed in~\cite{Porthos17}.

\subsection{Encoding the control-flow constraints}
\label{ch:port:enc:cf}

The control-flow of a program is represented in the \textit{control-flow graph}.

The control-flow graph is a directed acyclic connected graph with single source and multiple sink nodes, obtained by the \textit{loop unrolling} described in Section~\ref{ch:impl:xtoy:unrolling}.%TODO: footnote why multiple sinks
In control-flow graph, there are two types of edges: \textit{primary edges} that denote unconditional jumps or if-true-transitions (pictured with solid lines), and \textit{alternative edges} that denote if-false-transitions (pictured with dotted lines). Each node on graph can have either one successor (primary) or two successors (primary and alternative); only computation events can serve as a branching point). However, each merge node can have any positive number of predecessors, where each edge may be either primary or alternative.

While working on the \tool{mousquetaires}, we applied some optimisations of the encoding scheme for the control-flow in order to decrease the size of result formula. 
The \tool{PORTHOS} tool encodes the \textit{instructions}, which have recursive nature, in a recursive manner. This means, it adds synthetic composite instructions for both linear (sequential) and non-linear (branching) instructions. 
For instance, control-flow of the sequential instruction \m{i_1 := i_2; i_3} is encoded as 
\m{\phi_{CF}(i_2;i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \land cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)}, 
and control-flow of the branching instruction \m{i_1 := \{\mathtt{if} \, textbf \, \mathtt{then} \, i_2 \, \mathtt{else} \, i_3\}} is encoded as 
\m{\phi_{CF}(\mathtt{if} \, textbf \, \mathtt{then} \, i_2 \, \mathtt{else} \, i_3) = (cf_{i_1} \Leftrightarrow (cf_{i_2} \lor cf_{i_3})) \land \phi_{CF}(i_2) \land \phi_{CF}(i_3)}.

In contrast, the \tool{mousquetaires} firstly compiles the recursive high-level code into the linear low-level event-based representation,
% (similar to an assembly language)
that is then encoded into SMT formula. The encoding of branching nodes depends on the \textit{guards}, the value of conditional variable on the branching state, which in turn is encoded as data-flow constraint (see further in current Chapter).
%linearly as it shown in

Let \m{\fx : \mathbb{E} \rightarrow \{0,1\}} be the predicate that signifies the fact that the event has been e\textbf{x}ecuted (and, consequently, has changed the state of the system). 
Let \m{\fv : \mathbb{C} \rightarrow \mathbb{N}} be the function that returns the value of the computation event that will be computed once the event is executed (strictly speaking, it retuns the \textit{set} of values determined by the \texttt{rf}-relation; see Chapter~? for the relations encoding%TODO
). We distinguish the function \m{\fv_p : \mathbb{C_p} \rightarrow \{0,1\}} that evaluates the predicative computation event. In the result formula, all symbols $\fx(e)$ and $\fv(e)$ are encoded as boolean variables.

Consider the following possible mutual arrangement of nodes in a control-flow grpah:

\begin{figure}[H]
\centering
\begin{subfigure}[b]{0.3\textwidth}
\makebox[\textwidth]{
  \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
  \node[c] (1) [] {$e_1$};
  \node[c] (2) [below of=1] {$e_2$};
  \path[->]
  (1) edge [] node {} (2)
  ;
  \end{tikzpicture}
}
\caption{Sequential events}
\label{encode:cf:seq}
\end{subfigure}
~
\begin{subfigure}[b]{0.35\textwidth}
\makebox[\textwidth]{
  \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
  \node[c] (1) [] {$e_1$};
  \node[c] (2) [below left of=1] {$e_2$};
  \node[c] (3) [below right of=1] {$e_3$};
  \path[->]
  (1) edge [] node {} (2)
  (1) edge [dotted] node {} (3)
  ;
  \end{tikzpicture}
  }
\caption{Conditional branching}
\label{encode:cf:br}
\end{subfigure}
~
\begin{subfigure}[b]{0.25\textwidth}
\makebox[\textwidth]{
  \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
  \node[c] (i) [] {$e_i$};
  \node[c,draw=none] (ii) [right=0.2cm of i] {};
  \node[c,draw=none] (iii) [right=0.2cm of ii] {$...$};
  \node[c,draw=none] (iv) [right=0.2cm of iii] {};
  \node[c] (v) [right=0.2cm of iv] {$e_{i+j}$};
  \node[c] (k) [below of=iii] {$e_k$};
  \path[->]
  (i) edge [] node {} (k)
  (ii) edge [] node {} (k)
  (iii) edge [dotted] node {} (k)
  (iv) edge [dotted] node {} (k)
  (v) edge [] node {} (k)
  ;
  \end{tikzpicture}
}
\caption{Branch merging}
\label{encode:cf:merge}
\end{subfigure}
\caption{Linear and non-linear cases of control-flow graph}
\label{encode:cf}
\end{figure}

For listed cases, below we propose the encoding scheme that uniquely encodes each node of graph and allows to encode partially executed program.
The Equation~\ref{enc:cf:seq} for encoding the sequential control-flow represented on Figure~\ref{encode:cf:seq} reflects the fact that the event $e_2$ can be executed iff the event $e_1$ was executed. The Equation~\ref{enc:cf:br} for encoding the branching control-flow depicted on Figure~\ref{encode:cf:br} allows only following executions: $\{\emptyset, (e_1), (e_1 \rightarrow e_2), (e_1 \rightarrow e3) \}$. In encoding~\ref{enc:cf:merge} of the merge-point represented on Figure~\ref{encode:cf:merge}, the event $e_k$ is executed if either of its predecessors was executed, regardless of type of the transition.

\begin{align}
%\phi_{CF_{(e_1;e_2)}}     =
%\phi_{CF_{(e_1?e_2:e_3)}} =
\phi_{CF_{seq}}   = \ & \fx(e_2) \rightarrow \fx(e_1) \label{enc:cf:seq} \\
\phi_{CF_{br}}    = \ & [\fx(e_2) \rightarrow \fx(e_1)] \land [\fx(e_2) \rightarrow \fv(e_1)] \land \nonumber \\
				  & [\fx(e_3) \rightarrow \fx(e_1)] \land [\fx(e_3) \rightarrow \lnot\fv(e_1)] \land \nonumber \\
				  & \lnot [\fx(e_2) \land \fx(e_3)]  \label{enc:cf:br} \\
\phi_{CF_{merge}} = \ & \fx(e_k) \rightarrow (\bigvee\limits_{e_p \in \mathtt{pred}(e_k)}^{} \fx(e_p)) \label{enc:cf:merge}
\end{align}

\subsection{Encoding the data-flow constraints}
\label{ch:port:enc:df}

%As it was discussed in ...

% SINGLE-ASSIGNMENT

\subsection{Encoding the memory model constraints}
\label{ch:port:enc:wmm}

### 2018-04-02 1_wmm.tex
Add some words on control-flow encoding.

The event-based program model represents the directed graph (\textit{event-graph}), where vertices represent \textit{events}, and edges represent \textit{relations} over the events. An event is something, that, after being executed, changes the state of an abstract machine executing the concurrent program. 
An \textit{execution} (trace, run) of a given program is an ordered set of events.
%associated with the instructions of the program
%the path in the event-graph. 
An execution is considered to be \textit{valid} if the memory events follow a single global timeline, i.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}. An execution to be checked on validity is called the \textit{candidate execution}.

@@

A \textit{memory event} $e_m \in \mathbb{E}$ represents the fact of access to the memory. Since memory is the crucial low-level resource shared by multiple processes, most relations are defined over memory events. 
The processes can access a shared memory location (denoted by~$l_i$, for \textit{location}), or a local one (denoted by~$r_i$, for \textit{register}). A memory event is specified by its direction with respect to the shared variable, its location~$\texttt{loc}(e_m)$, its processor label~$\texttt{proc}(e_m)$, and a unique event label~$\texttt{id}(e_m)$~\cite{alglave2010shared}. 
%\texttt{load} for read the value of a shared-memory location, or \texttt{store} for write, or neither of them if both locations are local
The set of memory events $\mathbb{M}$ is devided into write events $\mathbb{W}$ (that write values to shared-memory locations) and read events $\mathbb{R}$ (that read values stored in shared-memory locations).
We add a restriction that each memory event uses at most one shared location, so that the write event $e_m = write(l_1, l_2)$ that writes value from shared location $l_2$ to the shared location $l_1$ is represented as two consequent events $e_m'~=~\texttt{load}(r_1 \leftarrow l_2); \ e_m''~=~\texttt{store}(l_1 \leftarrow r_1)$.

A \textit{computation event} $e_c \in \mathbb{C} \subseteq \mathbb{E}$, represents a low-level assembly computation operation performed solely on local-memory arguments. An example of computation event may be the event $e_c = r_3 \leftarrow add(r_1, r_2)$ that writes the sum of values stored in registers $r_1$ and $r_2$ to the register $r_3$. The \textit{control-flow} instructions (conditional and unconditional jumps) are encoded to the model directly, without additional events, as the $po$-relation (for \textit{program order}; see Chapter~\ref{ch:wmm:model:relations} for detailed definition of relations).

The third class of events is \textit{barrier events}, events caused by the synchronisation instructions (called \textit{fences}). Barrier events do not perform any computation or memory value transfer, instead, they add new relations to the model that restrict the set of allowed behaviours. Technically, a fence may either serve as a synchronisation barrier, or flush local memory caches, etc.

@@

The basic relation in the event-based program model is the \texttt{po}-relation $\subset~\mathbb{E}~\times~\mathbb{E}$ (\textit{program-order}), which represents the total order of memory events \textit{within single process}, which never relates events from different processes. Thus, if a program specifies the memory instruction $i_2$ to follow immideately the memory instruction $i_2$, then there exist an edge $e_1 \xrightarrow{\texttt{po}} e_2$ in the event-graph where event $e_1$ is caused by the instruction $i_1$ and $e_2$ is caused by the instruction $i_2$. This relation encodes the control-flow of the program into the event-graph.
%Some new relations may be acquired : dp, po-loc

@@

>>>>

The event-based program model represents the directed graph (\textit{event-graph}), where vertices represent \textit{events}, and edges represent \textit{relations} over the events. An event is something, that, after being executed, changes the state of an abstract machine executing the concurrent program. 
An \textit{execution} (trace, run) of a given program is an ordered set of events. The order of events in particular execution is denoted as `$\rightarrow$', an empty execution is denoted as $\emptyset$.
%associated with the instructions of the program
%the path in the event-graph. 
An execution is considered to be \textit{valid} if the memory events follow a single global timeline, textit.e., can be embedded in a single partial order allowed by the memory model restrictions~\cite{alglave2010shared}. An execution to be checked on validity is called the \textit{candidate execution}.

@@

A \textit{memory event} $e_m \in \mathbb{E}$ represents the fact of access to the memory. Since memory is the crucial low-level resource shared by multiple processes, most relations are defined over memory events. 
The processes can access a shared memory location (denoted by~$l_i$, for \textit{location}), or a local one (denoted by~$r_i$, for \textit{register}). A memory event is specified by its direction with respect to the shared variable, its location~$\mathtt{loc}(e_m)$, its processor label~$\mathtt{proc}(e_m)$, and a unique event label~$\mathtt{id}(e_m)$~\cite{alglave2010shared}. 
%\texttt{load} for read the value of a shared-memory location, or \texttt{store} for write, or neither of them if both locations are local
The set of memory events $\mathbb{M}$ is devided into write events $\mathbb{W}$ (that write values to shared-memory locations) and read events $\mathbb{R}$ (that read values stored in shared-memory locations).
We add a restriction that each memory event uses at most one shared location, so that the write event $e_m = write(l_1, l_2)$ that writes value from shared location $l_2$ to the shared location $l_1$ is represented as two consequent events $e_m'~=~\mathtt{load}(r_1 \leftarrow l_2); \ e_m''~=~\mathtt{store}(l_1 \leftarrow r_1)$.

A \textit{computation event} $e_c \in \mathbb{C} \subseteq \mathbb{E}$, represents a low-level assembly computation operation performed solely on local-memory arguments. An example of computation event may be the event $e_c = r_1 \leftarrow add(r_2, 1)$ that writes the sum of values stored in register $r_2$ and constant $1$ (emulated as a register as well) to the register $r_1$. For modelling branching statements, we distinguish the set $\mathbb{C_{p}} \subseteq \mathbb{C}$ of \textit{predicative} computation events, that are evaluated as a boolean value.%TODO: TAUTOLOGY HERE.

%The \textit{control-flow} instructions (conditional and unconditional jumps) are encoded into the model directly, without additional events, as the $po$-relation (for \textit{program order}; see Chapter~\ref{ch:wmm:model:relations} for detailed definition of relations).

The third class of events is \textit{barrier events}, events caused by the synchronisation instructions (called \textit{fences}). Barrier events do not perform any computation or memory value transfer, instead, they add new relations to the program model that restrict the set of allowed behaviours. Technically, a fence may be represented either as a synchronisation barrier, or a flush local memory caches, etc.

@@

The basic relation in the event-based program model is the \texttt{po}-relation $\subset~\mathbb{E}~\times~\mathbb{E}$ (\textit{program-order}), which represents the total order of memory events \textit{within single process}, which never relates events from different processes. Thus, if a program specifies the memory instruction $i_2$ to follow immideately the memory instruction $i_2$, then there exist an edge $e_1 \xrightarrow{\mathtt{po}} e_2$ in the event-graph where event $e_1$ is caused by the instruction $i_1$ and $e_2$ is caused by the instruction $i_2$. This relation encodes the control-flow of the program into the event-graph.
%Some new relations may be acquired : dp, po-loc

@@

// TODO: a bit more relations here

### 2018-04-02 2_impl.tex
Forgotten changes for release 14.03.2018.
Add some words on control-flow encoding.

\chapter{Implementation}
\label{section:impl}

@@

\section{Program Requirements}
\label{section:impl:requirements}

@@

\section{Program Components}
\label{section:impl:comp}
Big view

\section{C11 to YTree parser}
\label{section:impl:ytree}

@@

\section{YTree to XGraph event converter}
\label{section:impl:comp:xgraph}

@@

\subsection{Loop unrolling}
The original program encoded into the \texttt{XGraph} represents a \textit{flow graph}, a connected cyclic directed graph with single source node \texttt{[ENTRY]} (usually for convenience all leaves are connected to the sink node \texttt{[EXIT]}). The cycles are caused by low-level jump instructions, obtained from non-linear high-level control-flow statements (such as \texttt{while}, \texttt{do-while}, \texttt{for}, etc.). However, the cyclic flow graph cannot be encoded into SMT formula since ...
//TODO:REFERENCE.%TODO

\begin{figure}
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
\node[c] (1) [] {$1$};
\node[c] (2) [below left of=1] {$2$};
\node[c] (3) [below right of=1] {$3$};
\node[c] (4) [below of=3] {$4$};
\path[->]
(1) edge [] node {} (2)
(2) edge [bend left=50,dotted] node {} (1)
(1) edge [] node {} (3)
(3) edge [] node {} (4)
(4) edge [bend right=80,dotted] node {} (1)
;
\node[draw=none] (impl) [right=3cm of 3] {$\overset{k = 6}{\longmapsto}$};
;
\node[c] (21) [right=3cm of impl] {$2_1$};
\node[c] (11) [above right=1cm and 1cm of 21]{$1_1$};
\node[c] (31) [below right=1cm and 1cm of 11] {$3_1$};
\node[c] (41) [below of=31] {$4_1$};
\node[c] (12) [below of=21] {$1_2$};
\node[c] (22) [below left=1cm and 1cm of 12] {$2_2$};
\node[c] (32) [below right=1cm and 1cm of 12] {$3_2$};
\node[c] (13) [below of=41] {$1_3$};
\node[c] (14) [below of=22] {$1_4$};
\node[c] (43) [below of=32] {$4_3$};
\node[c] (23) [below of=13] {$2_3$};
\node[c] (33) [below of=13] {$3_3$};
\node[c] (24) [below left=1cm and 1cm of 14] {$2_4$};
\node[c] (34) [below right=1cm and 1cm of 14] {$3_4$};
\node[c] (15) [below right=1cm and -0.3cm of 43] {$1_5$};
\node[c] (44) [below of=33] {$4_4$};
\node[c] (6) [below left=1cm and 1cm of 15] {$6$};
\node[] (11k) [right=3.2cm of 11] {$(k = 1)$};
\node[] (31k) [right=1.7cm of 31] {$(k = 2)$};
\node[] (41k) [right=1.7cm of 41] {$(k = 3)$};
\node[] (13k) [right=1.7cm of 13] {$(k = 4)$};
\node[] (33k) [right=1.7cm of 33] {$(k = 5)$};
\node[] (44k) [right=1.7cm of 44] {$(k = 6)$};
\path[->]
(11) edge [] node {} (21)
(11) edge [] node {} (31)
(31) edge [] node {} (41)
(21) edge [] node {} (12)
(12) edge [bend right=10] node {} (22)
(12) edge [bend left=10] node {} (32)
(41) edge [] node {} (13)
(22) edge [] node {} (14)
(32) edge [] node {} (43)
(13) edge [] node {} (23)
(13) edge [] node {} (33)
(14) edge [bend right=10] node {} (24)
(14) edge [bend left=10] node {} (34)
(43) edge [] node {} (15)
(33) edge [bend right=10] node {} (15)
(33) edge [bend left=10] node {} (44)
(24) edge [bend right=20] node {} (6)
(34) edge [] node {} (6)
(15) edge [] node {} (6)
(44) edge [bend left=20] node {} (6)
;
\end{tikzpicture}
\label{fig:loop-unwind}
\caption{Example of the flow graph from the Figure~\ref{fig:merged-loop}, unwinded up to the bound $k = 6$} 
\end{figure}

\section{XGraph to ZFormula (SMT) encoder}
\label{section:impl:comp:zformula}

>>>>

\chapter{Implementation}
\label{ch:impl}

@@

\section{Program Requirements}
\label{ch:impl:requirements}

@@

\section{Program Components}
\label{ch:impl:comp}
Big view

\section{C11 to YTree parser}
\label{ch:impl:ytree}

@@

\section{YTree to XGraph event converter}
\label{ch:impl:xtoy}

@@

\subsection{Loop unrolling}
\label{ch:impl:xtoy:unrolling}
% If the analysing program contains a loop, then its control-flow graph represents the cyclic directed graph. Although, in order to be able to encode such a cyclic structure as a boolean formula, this graph needs to be  
% In order to be

The original program encoded into the \texttt{XGraph} represents a \textit{flow graph}, a connected cyclic directed graph with single source node \texttt{[ENTRY]} (usually for convenience all leaves are connected to the sink node \texttt{[EXIT]}). The cycles are caused by low-level jump instructions, obtained from non-linear high-level control-flow statements (such as \texttt{while}, \texttt{do-while}, \texttt{for}, etc.). However, the cyclic flow graph cannot be encoded into SMT formula since ...
//TODO:REFERENCE.%TODO

\begin{figure}
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.5cm,semithick]
\node[c] (1) [] {$1$};
\node[c] (2) [below left of=1] {$2$};
\node[c] (3) [below right of=1] {$3$};
\node[c] (4) [below of=3] {$4$};
\path[->]
(1) edge [] node {} (2)
(2) edge [bend left=50,dotted] node {} (1)
(1) edge [] node {} (3)
(3) edge [] node {} (4)
(4) edge [bend right=80,dotted] node {} (1)
;
\node[draw=none] (impl) [right=3cm of 3] {$\overset{k = 6}{\longmapsto}$};
;
\node[c] (21) [right=3cm of impl] {$2_1$};
\node[c] (11) [above right=1cm and 1cm of 21]{$1_1$};
\node[c] (31) [below right=1cm and 1cm of 11] {$3_1$};
\node[c] (41) [below of=31] {$4_1$};
\node[c] (12) [below of=21] {$1_2$};
\node[c] (22) [below left=1cm and 1cm of 12] {$2_2$};
\node[c] (32) [below right=1cm and 1cm of 12] {$3_2$};
\node[c] (13) [below of=41] {$1_3$};
\node[c] (14) [below of=22] {$1_4$};
\node[c] (43) [below of=32] {$4_3$};
\node[c] (23) [below of=13] {$2_3$};
\node[c] (33) [below of=13] {$3_3$};
\node[c] (24) [below left=1cm and 1cm of 14] {$2_4$};
\node[c] (34) [below right=1cm and 1cm of 14] {$3_4$};
\node[c] (15) [below right=1cm and -0.3cm of 43] {$1_5$};
\node[c] (44) [below of=33] {$4_4$};
\node[c] (6) [below left=1cm and 1cm of 15] {$6$};
\node[] (11k) [right=3.2cm of 11] {$(k = 1)$};
\node[] (31k) [right=1.7cm of 31] {$(k = 2)$};
\node[] (41k) [right=1.7cm of 41] {$(k = 3)$};
\node[] (13k) [right=1.7cm of 13] {$(k = 4)$};
\node[] (33k) [right=1.7cm of 33] {$(k = 5)$};
\node[] (44k) [right=1.7cm of 44] {$(k = 6)$};
\path[->]
(11) edge [] node {} (21)
(11) edge [] node {} (31)
(31) edge [] node {} (41)
(21) edge [] node {} (12)
(12) edge [bend right=10] node {} (22)
(12) edge [bend left=10] node {} (32)
(41) edge [] node {} (13)
(22) edge [] node {} (14)
(32) edge [] node {} (43)
(13) edge [] node {} (23)
(13) edge [] node {} (33)
(14) edge [bend right=10] node {} (24)
(14) edge [bend left=10] node {} (34)
(43) edge [] node {} (15)
(33) edge [bend right=10] node {} (15)
(33) edge [bend left=10] node {} (44)
(24) edge [bend right=20] node {} (6)
(34) edge [] node {} (6)
(15) edge [] node {} (6)
(44) edge [bend left=20] node {} (6)
;
\end{tikzpicture}
\label{fig:loop-unwind}
\caption{Example of the flow graph from the Figure~\ref{fig:merged-loop.....}, unwinded up to the bound $k = 6$} 
\end{figure}

\section{XGraph to ZFormula (SMT) encoder}
\label{ch:impl:comp:zformula}
